---
output: 
  pdf_document:
    includes:
      in_header: imports.sty
      before_body: header.sty
      toc: true
      toc_depth: 3
urlcolor: blue
---
<!--the above imports all packages for this file as well as defines the table of contents and cover page for the project.-->
```{r, include=FALSE}
#### Generate a Function to Ensure a Package is Installed ###
EnsurePackage<-function(x){
  x<-as.character(x)
  if (!require(x,character.only=TRUE)){
    install.packages(pkgs=x, repos="http://cran.r-project.org")
    require(x, character.only=TRUE)
  }
}
###########Import All Required Packages#######################
EnsurePackage("compare")
EnsurePackage("ggplot2")
EnsurePackage("ggmap")
EnsurePackage("gridExtra")
EnsurePackage("rgeos")
EnsurePackage("maptools")
EnsurePackage("RJSONIO")
EnsurePackage("sqldf")
```
<!--this is the main body of the project assignment-->
#1 Introduction
As a once long-term resident of the great state of New York. The author aims to learn more about its most populous city through this project. The City of New York, New York (NYC) has generated an Open Data imitative in order to be fair and open with its constituents. As the author may be taking jobs in the city as early as 2020, there must be a way to determine the dangerous areas of the city from the crime data. In order to do this the student has referenced the New York Police Department (NYPD) public facing records for shooting crime incidents for the period of 2013 through 2018. Throughout this project chunks of **R** code will be displayed inline as a reference to better understand the process.

***
#2 Business Questions
This project will attempt to answer the following three questions:

*	Does time of day and location have a direct correlation to a shooting event?
* Has any part of the city been consistently high in shooting incidents?
* Is any one particular demographic group at high risk ?

As the purpose of this project is to determine the dangerous parts of NYC. Understanding the time of day that most shootings occur as well as which areas of the city these incidents occur in will be key to a new resident choosing their domicile. Additionally, as a bi-product of the information we can generate information on at risk demographics. While project will not predict the housing market prices or the cost of living, it will give a reader a reasonable glimpse at the public order of the city.   

***
#3 Data Acquisition, Cleansing Transformation, Munging

##3.1 Problem Definition

The purpose of this project is to generate actionable data for a potential future resident of NYC. Actionable data is defined by this project as tables, charts, and modeling to provide graphical information for the reader to understand where the shooting hotspots in the last five years of shooting incidents in NYC. The data from [New York Police Department](https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Historic-/833y-fsy8 "NYPD Shooting Incident Data (Historic)") is assigned several flags that can better help define the demographics of the project. There are some key attributes within this dataset that can be used to generate actionable data. 

Exmaple Attributes:

* Ages of Accused & Victims
* Sex of Accused & Victims
* Location of the Crime
* Time and Date of the Crime

##3.2 Data Acquisition 
The **NYPD Shooting Incident Data**, which is available freely from the [NYC Open Data website](https://opendata.cityofnewyork.us/ "NYC Open Data") contains one set of data. This project will utalize the powerful data science oriented programing language of **R** to analyze this data. Upon intial review of the information provided to the author via the NYPD. Reading the source documentation for this file it was determined that the best method to import would be a Comma Seperated Values (CSV). Addtionally the document retrieved should contain a total of 6,407 rows with 18 attributes for a total of 115,326 data points. In order to start processing the data it was first imported into the system and stored in the variable rawCSV.

```{r, results=FALSE}
urlToImport <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
rawCSV <- data.frame((read.csv(urlToImport, stringsAsFactors = FALSE)))
```

Generating the vector of rawCSV, we created a data frame with the contents of the download csv button from our selected source. Wtih this data now stored within our project, a data scieentist must confirm that the information has arrived correct and that it is able to be used within the project.

##3.3 Data Clensing Process
Although the data retrived is in relativly good shape there are some peculiarities that need to be addressed during the porocess in order to properly utalize this data within the project. The first piece that could be easily recognized just in a look at the data is that alot of the information is in capital letters. Due to several proceses within this project requiring the lowercase data, a to lower operation was required to be conducted on a few fields. 

```{r, results=FALSE}
cleanCSV <- rawCSV
colnames(cleanCSV) <- tolower(colnames(cleanCSV))
cleanCSV$boro<-tolower(cleanCSV$boro)
cleanCSV$location_desc<-tolower(cleanCSV$location_desc)
cleanCSV$perp_sex<-tolower(cleanCSV$perp_sex)
cleanCSV$perp_race<-tolower(cleanCSV$perp_race)
cleanCSV$vic_sex<-tolower(cleanCSV$vic_sex)
cleanCSV$vic_race<-tolower(cleanCSV$vic_race)
```

The next issue with the intial importing of the CSV file is that the file has not properly converted all the fields to their proper data types. Upon pulling the data frames structure we can see that some fields that should be numbers are characters, as well as fields that should be dates or times are also treated as a character field. 

```{r, results=FALSE}
str(cleanCSV)
```

In order to correct these issues, the first column to be targeted was that of the occur_date. From the above structure you can see that this is shown as both a factor and additionally it contained both an unformated date as well as some erronious informationn inserted somewhere along the way during the csv file creation process. 

```{r, results=FALSE}
removeTime <- function(inputVector){
  inputVector <-gsub(" 12:00:00 AM","",inputVector)
  inputVector <- gsub(" ", "", inputVector)
  inputVector <- as.Date(inputVector, format='%m/%d/%Y')
  return(inputVector)
}
cleanCSV$occur_date<-removeTime(cleanCSV$occur_date)
```

The column titled location_desc is filled with limited information for the purposes of this report. When it is filled with data the information is not consistant or actionable. Additionally, there is a column to indicate what type of officer was involved in this incident. These columns were removed from the data frame and than verified that it was indeed removed. 

```{r, results=FALSE}
cleanCSV<-cleanCSV[,-6:-7]
colnames(cleanCSV)
```

Finally, after all of this cleaning has been completed only complete rows should remain. This will be rows that do not have any blank values that cannot be used within this report. 

```{r, results=FALSE}
cleanCSV<-cleanCSV[complete.cases(cleanCSV),]
```

After performing all of these operations to prepare the data for use within the report the data contains 6407 rows and 17 attributes to perform work on for a total of 108,919 data points. 

##3.4 Data Dictionary
The following tables displays the list of data items used within this report. 

```{r, echo=FALSE}
varName<-colnames(cleanCSV)
varType<-c("int", 
           "date", 
           "chr", 
           "chr", 
           "int", 
           "chr", 
           "chr", 
           "chr", 
           "chr", 
           "chr", 
           "chr", 
           "chr", 
           "int", 
           "int", 
           "num",
           "num")
varDesc<-c("Incidient number, duplicates are victims",
           "Indicates the date of occurance",
           "Indicates the time of the incident",
           "Indicates which borrough the even occured in",
           "Indicates which police department responded",
           "Indicates if victim died",
           "If known, age range of perpetrator",
           "If known, sex of perpetrator",
           "If known race of perpetrator ",
           "Victims Age Range","Victims Sex",
           "Victims Race",
           "X grid coordinate for event",
           "Y grid coordinate for event",
           "Lattitude of event",
           "Longitude of event")
dataDict<-data.frame(varName,varType,varDesc)
colnames(dataDict)<-c("Variable Name", "Variable Type", "Variable Description")

knitr::kable(dataDict, "latex", booktabs=T)
```

***
#4 Descriptive Statistics

##4.1 Summary Statistics
Prior to doing any work with the data, the author wanted to understand the data that they were using. Using the summary command they were able to get the minimum and max values, as well as the mean and median values. As the incident key is an auto generated number for each incident it is not included in these tables. 
```{r, echo=FALSE}
knitr::kable(summary(cleanCSV[2:6]), "latex", booktabs=T)
knitr::kable(summary(cleanCSV[7:11]), "latex", booktabs=T)
knitr::kable(summary(cleanCSV[12:16]), "latex", booktabs=T)
```

##4.2 Data Structure
In order to process and manipulate the data within the project, we must get a firm grasp for what each fields data type is in **R**. This is reflected by using the structure command, this command gives a snapshot of all the columns from a dataset and its contents. 
```{r, echo=FALSE}
str(cleanCSV)
```

##4.3 Graphs, Charts, Tables
The first part of generating charts and graphs is to generate the data into easier to read tables for the purpose of genarting graphs to be read. 
```{r results=FALSE}
deaths <- sqldf("SELECT *
                FROM cleanCSV
                WHERE statistical_murder_flag IN ('true')")
injuries <- sqldf("SELECT *
                FROM cleanCSV
                  WHERE statistical_murder_flag NOT IN ('true')")

incidentsByLocation <- sqldf("SELECT DISTINCT incident_key, x_coord_cd, 
                            y_coord_cd, latitude, longitude
                             FROM cleanCSV")

```

The following graph shows the number of shooting incidents broken down by 30 day periods. Although this does not overlap well for months longer or shorter it is a baseline that sits well over the whole year. 

```{r echo=FALSE}
hist_events <- ggplot(data=cleanCSV, aes(x=occur_date)) + geom_histogram(stat = "bin", binwidth = 30,fill="Red",color='Black') + theme(legend.position = "none")
hist_events <- hist_events + ggtitle("Number of Shooting Incidents Per Month")+xlab("Date of Shooting")+ylab("Number of Shooting Victims")
hist_events
```

The next chart will be the number of murders over the same period of time. Although the shooting events are important. These individuals have lived through this ordeal as far as the NYPD was aware at the time of this innformations publication. The author had difficulty getting the **R** to display the TRUE flag as an instance of one murder being taken place. Even after converting the values to 1 or 0 the chart continued to display as true or false values. 

```{r, echo=FALSE}
deaths$statistical_murder_flag<-as.integer(as.logical(deaths$statistical_murder_flag))
line_deaths <- ggplot(data=cleanCSV, aes(x=occur_date, y=statistical_murder_flag))+geom_line()
line_deaths <- line_deaths + ggtitle("Number of Murders")+xlab("Date of Shooting")+ylab("Number of Murders")
line_deaths
```

***
#5 Modeling Techniques

The author had issues converting the data into values that could be understood by the linear modeling techniques shown in class. Below is the code that they attempted to introduce for the murder being related to the borrogh to which the shooting occured. 
```{r, eval=FALSE}
correlations <-  ggplot(data=cleanCSV, aes(y=statistical_murder_flag, x=boro)) + geom_point()
lm1 <- lm(formula = statistical_murder_flag~boro, cleanCSV)
correlations
abline(lm1)
```

***
#6 Summary

This report was designed to take information published by the NYPD and return from it information that a person moving to the city could make use of during the 2019 year. Reading from the data presetned you can see that there were 6407 shootign injuries between the years of 2013 and 2018 in New York City. From these shootings there were 1147 declared as murder. This means that `r (1147/6407)*100` % of all shootings resulted in a murder. Additionnally if you take a look at the numbers of incidents that have taken place from 2013 until 2018 you will see that the data has stayed roughly constant with a spike in events in early 2018. That is to say that the risk of you getting shot in NYC has stayed cosntant and that if the population of NYC is millions of people, your chances of getting shot are very slim. This report would recommend that you do move to NYC and good luck on your future journey. 

***
#8 Appendix
##A.1 R Code

```{r, eval=FALSE}
#### Generate a Function to Ensure a Package is Installed ###
EnsurePackage<-function(x){
  x<-as.character(x)
  if (!require(x,character.only=TRUE)){
    install.packages(pkgs=x, repos="http://cran.r-project.org")
    require(x, character.only=TRUE)
  }
}
removeTime <- function(inputVector){
  inputVector <-gsub(" 12:00:00 AM","",inputVector)
  inputVector <- gsub(" ", "", inputVector)
  inputVector <- as.Date(inputVector, format='%m/%d/%Y')
  return(inputVector)
}

###########Import All Required Packages#######################
EnsurePackage("compare")
EnsurePackage("ggplot2")
EnsurePackage("ggmap")
EnsurePackage("gridExtra")
EnsurePackage("rgeos")
EnsurePackage("maptools")
EnsurePackage("RJSONIO")
EnsurePackage("sqldf")

#############Import Source Data for Project###################
urlToImport <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
rawCSV <- data.frame((read.csv(urlToImport, stringsAsFactors = FALSE)))

########################Pre-Processing Steps##################
cleanCSV <- rawCSV
colnames(cleanCSV) <- tolower(colnames(cleanCSV))
cleanCSV$boro<-tolower(cleanCSV$boro)
cleanCSV$location_desc<-tolower(cleanCSV$location_desc)
cleanCSV$perp_sex<-tolower(cleanCSV$perp_sex)
cleanCSV$perp_race<-tolower(cleanCSV$perp_race)
cleanCSV$vic_sex<-tolower(cleanCSV$vic_sex)
cleanCSV$vic_race<-tolower(cleanCSV$vic_race)


cleanCSV$occur_date<-removeTime(cleanCSV$occur_date)

cleanCSV<-cleanCSV[,-6:-7]
colnames(cleanCSV)

cleanCSV<-cleanCSV[complete.cases(cleanCSV),]

####################Creating Descriptive Statistics###########

knitr::kable(summary(cleanCSV[2:6]), "latex", booktabs=T)
knitr::kable(summary(cleanCSV[7:11]), "latex", booktabs=T)
knitr::kable(summary(cleanCSV[12:16]), "latex", booktabs=T)

str(cleanCSV)

#################Generating Graph Data########################

deaths <- sqldf("SELECT *
                FROM cleanCSV
                WHERE statistical_murder_flag IN ('true')")
injuries <- sqldf("SELECT *
                FROM cleanCSV
                  WHERE statistical_murder_flag NOT IN ('true')")

incidentsByLocation <- sqldf("SELECT DISTINCT incident_key, x_coord_cd, y_coord_cd, latitude, longitude
                             FROM cleanCSV")


########Create the Data Dictionary############################
varName<-colnames(cleanCSV)
varType<-c("int", 
           "date", 
           "chr", 
           "chr", 
           "int", 
           "chr", 
           "chr", 
           "chr", 
           "chr", 
           "chr", 
           "chr", 
           "chr", 
           "int", 
           "int", 
           "num",
           "num")
varDesc<-c("Incidient number, duplicates are victims",
           "Indicates the date of occurance",
           "Indicates the time of the incident",
           "Indicates which borrough the even occured in",
           "Indicates which police department responded",
           "Indicates if victim died",
           "If known, age range of perpetrator",
           "If known, sex of perpetrator",
           "If known race of perpetrator ",
           "Victims Age Range","Victims Sex",
           "Victims Race",
           "X grid coordinate for event",
           "Y grid coordinate for event",
           "Lattitude of event",
           "Longitude of event")
dataDict<-data.frame(varName,varType,varDesc)
colnames(dataDict)<-c("Variable Name", "Variable Type", "Variable Description")

knitr::kable(dataDict, "latex", booktabs=T)

########Create the map to plot on#############################

hist_events <- ggplot(data=cleanCSV, aes(x=occur_date)) + geom_histogram(stat = "bin", binwidth = 30,fill="Red",color='Black') + theme(legend.position = "none")
hist_events <- hist_events + ggtitle("Number of Shooting Incidents Per Month")+xlab("Date of Shooting")+ylab("Number of Shooting Victims")
hist_events

deaths$statistical_murder_flag<-as.integer(as.logical(deaths$statistical_murder_flag))
line_deaths <- ggplot(data=cleanCSV, aes(x=occur_date, y=statistical_murder_flag))+geom_line()
line_deaths <- line_deaths + ggtitle("Number of Shooting Incidents Per Month")+xlab("Date of Shooting")+ylab("Number of Shooting Victims")
line_deaths

##############Plot out a LM of the deaths to burrough########

correlations <-  ggplot(data=cleanCSV, aes(y=statistical_murder_flag, x=boro)) + geom_point()
lm1 <- lm(formula = statistical_murder_flag~boro, cleanCSV)
correlations
abline(lm1)
```