---
output: 
  pdf_document:
    includes:
      in_header: imports.sty
      before_body: header.sty
      toc: true
---
<!--the above imports all packages for this file as well as defines the table of contents and cover page for the project.-->
```{r, include=FALSE}
#### Generate a Function to Ensure a Package is Installed ###
EnsurePackage<-function(x){
  x<-as.character(x)
  if (!require(x,character.only=TRUE)){
    install.packages(pkgs=x, repos="http://cran.r-project.org")
    require(x, character.only=TRUE)
  }
}
###########Import All Required Packages#######################
EnsurePackage("ggplot2")
EnsurePackage("ggmap")
EnsurePackage("maptools")
EnsurePackage("RJSONIO")
```
<!--this is the main body of the project assignment-->
#1 Introduction
As a once long term resident of the great state of New York. The author aims to learn more about its most populious city through this project. The City of New York, New York has generated an Open Data iniative in order to be fair and open with its consituents. As the author may be taking jobs in the city as early as 2020, there must be a way to determine the dangerous areas of the city from the crime data. In order to do this the student has referenced the New York Police Department (NYPD) public facing records for shooting crime incidents for the period of 2013 through 2018. Throughout this project chunks of **R** code will be displayed inline as a reference to better understand the process. 

***
#2 Business Questions
This project will attempt to answer the following two questions:

*	Does time of day and location have a direct corilation to a shooting event?
* Has any part of the city consistantly been aprt of the 

Reason for this understanding

***
#3 Data Acquisition, Cleansing Transformation, Munging

##3.1 Problem Definition

##3.2 Data Acquisition 
  The data for this report was gathered from [New York Cities Open Data Project](https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Historic-/833y-fsy8 "NYC OpenData Project"). The initial setup for this file is a comma seperated values document with a total of 6,407 rows with 18 attributes for a total of 115,326 data points. 


```{r, eval=FALSE}
urlToImport <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
rawCSV <- data.frame((read.csv(urlToImport)))
```

##3.3 Data Clensing Process

##3.4 Data Dictionary

***
#4 Descriptive Statistics

##4.1 Summary Statistics

##4.2 Data Structure

##4.3 Re-shaping the Data

##4.4 Graphs, Charts, Tables

#5 Modeling Techniques

#6 Data Interpretation

#7 Summary

Actionable ideas or insights 

#8 Appendix
##A.1 R Code

```{r, eval=FALSE}
#### Generate a Function to Ensure a Package is Installed ###
EnsurePackage<-function(x){
  x<-as.character(x)
  if (!require(x,character.only=TRUE)){
    install.packages(pkgs=x, repos="http://cran.r-project.org")
    require(x, character.only=TRUE)
  }
}
###########Import All Required Packages#######################
EnsurePackage("ggplot2")
EnsurePackage("ggmap")
EnsurePackage("gridExtra")
EnsurePackage("maptools")
EnsurePackage("RJSONIO")

##############Gather and Load the Data########################
urlToImport <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
rawCSV <- data.frame((read.csv(urlToImport)))

########Create the Data Dictionary############################
varName<-c("list of names")
varType<-c("list of types")
varDesc<-c("list of descriptions")
dataDict<-data.frame(varName,varType,varDesc)
colnames(dataDict)<-c("Variable Name", "Variable Type", "Variable Description")
grid.table(dataDict)

########Create the map to plot on#############################
```
##A.2 Notes